---
title: "Demo Day" 
author: "Liz Masten"
date: "4/29/2020"
output: bookdown::pdf_document2
bibliography: bib2.bib
biblio-style: apalike 
link_citations: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(foreign)
library(tinytex)
library(bookdown)
library(bibtex)
library(yaml)
library(knitcitations)
library(magick)
library(memisc)
library(pscl)
library(MASS)
library(amen)
library(cluster)
library(lmtest)
library(plm)
library(sandwich)
library(stargazer)
library(huxtable)

```

# Abstract 

Findley, Piazza, and Young (2012) show that interstate rivalries are a positive predictor of transnational terrorist activity. The authors argue that terrorism is often a component of broader hostilities that can be emperically analyzed using a series of politicaly relevant directed dyads. I was able to successfully replicate all of their results. In my extension, I run dyadic quasipoisson models using R. This model is equally statistically significant with respect to rivalry, the main concern of the original paper, and thereby confirms that the authors' original findings are robust.  

# Introduction 

Findley, Piazza, and Young analyze transnational terrorism as a component of interstate rivalries, specifically focusing on the use of terrorism in proxy warfighting. Their main argument is that terrorist attacks are more likely to occur in the context of a rivalry between two states than in the absence of such a rivalry. They empirically test their hypothesis by analyzing “politically relevant directed state dyads.” They define politically relevant as states where “relationships of interest are at least possible” and utilize directed state dyads to indicate directionality of attacks (i.e. the state where the attack originated) in which there exists at least one major power [@findley2012games].

I was able to successfully replicate the authors' findings in this paper exactly. Note that the authors suffered from a data management error at the time of publication and subsequently published errata results. I based my analysis on the revised data and tables, which can be found in the Appendix. In an effort to not reinvent the wheel, and because there are no directly analagous R packages to translate the authors' original STATA code, I wrote and managed the project in R while running the replication in STATA. My extension is done in R. 

In my extension, I run four dyadic quasipoisson models using the same data as the original authors. These models are similarly statistically significant to the original results, and the statistical significance of the rivalry variable is equivalent to the original findings. In fact, my models report a stronger effect of rivalry than the orignal paper. These results confirm that the authors' original findings are robust.

This paper is laid out as follows: in the section below, I discuss relevant literature as it relates to the concept of interstate rivalries, state-sponsored terrorism, and the use of proxies. I then discuss the research design and analysis of Findley et al. (2012) and my successful replication of their results. Next, I conduct an extension which validates the authors' findings using dyadic quasipoisson models. Lastly, I discuss my results and the implications for future emperical research in the field of terrorism studies.^[All analysis for this paper is available in my GitHub: https://github.com/LizMas/1006_replication].

# Literature Review

States both use and suffer from terrorism. It is no secret that states support terrorist groups, actively or passively, willingly or unwillingly. States with strong and stable institutions may give groups direct aid including money, weapons, or logistic support. They may help with training and operations, back them diplomatically, assist organizing efforts, steer the group’s ideological direction, or provide sanctuary. States with weak capacity or institutional infighting may offer support in similar ways, but with the caveat that they couldn’t get rid of the group or its influence if they wanted to [@byman2010agents].    

Terrorist groups make appealing proxies because they specialize in activities that states are often unwilling or unable to do directly. They offer cover and plausible deniability in case an operation goes bad or is unpopular, allowing states to avoid directly committing an act of war against another state [@conrad2011interstate]. Terrorist proxies are also disproportionally effective when considering the amount of money, time, and personnel it takes to orchestrate an attack versus the capital, infrastructural, and psychological damage they can inflict. This disproportionality makes proxies especially attractive to weak states who have rivalries with stronger states (Iran prolifically uses proxies to counter Saudi, US, and Israeli influence in the Middle East). States may also delegate to terrorist groups to ensure the state’s preferences continue beyond the effective global or regional influence of that state [@byman2010agents]. Iranian support of nascent Hezbollah is a good case study of this. 

The often surreptitious nature of terrorist proxies makes analyzing this phenomenon difficult.  The empirical study of terrorism, as a subfield of International Relations, tries to make sense of it in two main ways: using a country-year method of analysis, in which variables are analyzed continuously, and dyadic analysis, in which countries are paired and models are estimated having dyadic measurements as the outcome variable. International Relations scholars have a strong preference for dyadic analysis when possible, but terrorism scholars have been somewhat reticent to employ it. Terrorism scholars have traditionally favored using a domestic social and institutional or micro-foundational levels of analysis [@cranmer2016critique]. 

Findley et al. (2012) were among the first scholars to use dyadic analysis in the field. For their experiment design, they employ a dyadic design to investigate the causal relationship between interstate rivalry and terrorism. They chose dyadic analysis because they deemed it “crucial for establishing causal chains between indicators and patterns of terrorism because it permits us to examine the interplay of both origin countries for terrorism—the countries from which terrorists hail—and target countries that experience terrorist attacks.” 

They use directed dyads, as opposed to nondirected dyads, because they argue that there may be “that certain aspects of the origin country—the country of national origin of the perpetrators of the attack—lead to greater attacks on the target country—the country where the attack occurred. Directed dyads are also appropriate because certain aspects of the target country may make them a more plausible target for actors from the origin country.” They also used only politically relevant dyads, which allows them to discount pairs of states which would not significantly interact in the international system. This keeps the number of observations from being deceivingly high, which would increase the likelihood of spurious statistical significance. 

# Replication 

I successfully replicated the results from Findley et al. (2012). The code for this replication and the tables it produces can be found in the Appendix. 

# Extension

```{r}

# data from errata 

#na.omit to get rid of 15,000 observations in joint democracy. This was necessary for the clustering function to run. 

data <- read.dta("updated_data.dta") %>% na.omit()

```

```{r}

#write clustering function for dyadic analysis. Thanks to Alice Xu for help with this: 

cluster_se <- function(model, cluster){
  if(nrow(model.matrix(model))!=length(cluster)){
    stop("check your data: cluster variable has different N than model")}
  M <- length(unique(cluster))
  N <- length(cluster)          
  K <- model$rank  
  dfc <- (M/(M - 1)) * ((N - 1)/(N - K))
  uj  <- apply(estfun(model), 2, function(x) tapply(x, cluster, sum));
  rcse.cov <- dfc * sandwich(model, meat = crossprod(uj)/N)
  return(coeftest(model, vcov. = rcse.cov))
}

```

```{r}

# NOTE: The authors did a negative binomial regression in STATA using nbreg. The R equivalent is glm.nb from the MASS package. However, glm.nb sets dispersion at 1 by default. Quasipoisson allows for the dispersion to be unrestricted by the model and rather determined from the data itself.   

# Model 1: STATA: 'nbreg terrorCounts rivalry jointDem1 logcapratio contiguity, nolog cluster(dyadid) dispersion(constant)' 

m1 <- glm(data = data, terrorCounts ~ rivalry + jointDem1 + logcapratio + contiguity, family = quasipoisson)

# STATA: 'nbreg terrorCounts2 rivalry jointDem1 logcapratio contiguity, nolog cluster(dyadid)  dispersion(constant)'  

m2 <- glm(data = data, terrorCounts2 ~ rivalry + jointDem1 + logcapratio + contiguity, family = quasipoisson) 

# STATA: 'nbreg terrorCounts rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, nolog cluster(dyadid) dispersion(constant)' 

m3 <- glm(data = data, terrorCounts ~ rivalry + jointDem1 + logcapratio + historyl1 + historyl2 + coldwar1 + conflict1 + conflict2 + contiguity + war1 + war2, family = quasipoisson) 

#STATA: 'nbreg terrorCounts2 rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, nolog cluster(dyadid)  dispersion(constant)' 

m4 <- glm(data = data, terrorCounts2 ~ rivalry + jointDem1 + logcapratio + historyl1 + historyl2 + coldwar1 + conflict1 + conflict2 + contiguity + war1 + war2, family = quasipoisson) 

```


```{r}

#run the function on the data. Equivalent of cluster(dyadid) from STATA:  

mod_cluster_1 <- cluster_se(m1, data$dyadid)

mod_wald_1 <- waldtest(m1)

mod_cluster_2 <- cluster_se(m2, data$dyadid)

mod_wald_2 <- waldtest(m2)

mod_cluster_3 <- cluster_se(m3, data$dyadid)

mod_wald_3 <- waldtest(m3)

mod_cluster_4 <- cluster_se(m4, data$dyadid)

mod_wald_4 <- waldtest(m4)

```

```{r ext_table, warning = FALSE, message = FALSE, echo = FALSE}


#ext_table <- stargazer(mod_cluster_1, mod_cluster_2, mod_cluster_3, mod_cluster_4, 
               #     type = 'latex',
                #    title = "Quasipoisson Clustered Models of Transnational Terrorist Activity per Findley et al (2012)")

#ext_table

#stargazer broke latex? 

print(mod_cluster_1)
print(mod_cluster_2)
print(mod_cluster_3)
print(mod_cluster_4)

``` 

```{r wald, results= 'asis', echo=FALSE, message=FALSE, warning=FALSE}

#wald_table <- stargazer(mod_wald_1, mod_wald_2, mod_wald_3, mod_wald_4, type = "latex")

#wald_table

```


```{r generated bibliography, echo=FALSE, message=FALSE, warning=FALSE}

cleanbib()
options("citation_format" = "pandoc")
read.bibtex(file = "bib2.bib")

#This is still a mess, but I can't for the life of me get the # off.
 
```
# Appendix 

Appendix A: Original Table 
```{r, warning = FALSE, echo = FALSE, error = FALSE}

# This is the orignal fpy table: 

img <- magick::image_read('og_table_main.png')
print(img)

```

Appendix B: Replication of results 

Note: This table reports non-robust standard errors.  
```{r, warning = FALSE, echo = FALSE, error = FALSE}

# this is my replication, using regular SE instead of Robust SE because I couldn't get STATA to report the robust SEs. They were the same though. 

img2 <- magick::image_read('stata_rep_main.png')
print(img2)

```


```{r, stata code}

#this is the STATA code for the replication. DO NOT try to run in R. 


#use "fpy replication data jop 2012.dta"
#set more off

#/*NOTE: original code from fpy, adjusted for the replication*/ 

#******************
#* Table 1        *
#******************

#/* Table 1, Model 1*/

#nbreg terrorCounts rivalry jointDem1 logcapratio contiguity, nolog cluster(dyadid) dispersion(constant) robust

#/*store table, do for each model*/ 

#estimates store m1

#/*Table 1, Model 2 */

#nbreg terrorCounts2 rivalry jointDem1 logcapratio contiguity, nolog cluster(dyadid)  dispersion(constant) robust

#estimates store m2

#/*Table 1, Model 3 */
#nbreg terrorCounts rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, nolog cluster(dyadid) dispersion(constant) robust

#estimates store m3

#/*Table 1, Model 4 */
#nbreg terrorCounts2 rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, nolog cluster(dyadid)  dispersion(constant) robust

#estimates store m4

#/*Output table*/ 

#estout m1 m2 m3 m4, cells(b se(par))

#********************
#* Table 2          *
#********************

#/* Table 2, Model 5 */

#zinb terrorCounts rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, inflate(rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2) cluster(dyadid) nolog robust

#estimates store m5

#/* Table 2, Model 6 */
#zinb terrorCounts2 rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, inflate(rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2) cluster(dyadid) nolog robust

#estimates store m6

#/* Table 2, Model 7 */
#/* Note: convergence issues require a different maximization technique (bfgs) */
#zinb terrorCounts rivalry jointDem1 logcapratio historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, inflate(jointDem1) cluster(dyadid)  technique(bfgs) nolog robust

#estimates store m7

#/* Table 2, Model 8 */
#zinb terrorCounts2 rivalry jointDem1 logcapratio  historyl1 historyl2 coldwar1 conflict1 conflict2 contiguity war1 war2, inflate(jointDem1) cluster(dyadid) nolog robust

#estimates store m8

#/*Output table*/ 

#estout m5 m6 m7 m8, cells(c se(par)) 

```
# References
